---
title: "Final Project"
author: "Hongshuo Zhou"
output: 
  pdf_document:
    toc: true
    number_sections: true
    # theme: united
---
\newpage
```{r load-data, echo=FALSE, include=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

library(ggplot2)
library(leaflet)
library(readxl)
#install.packages("DT")
library(patchwork)
library(gridExtra)
library(stringr)
library(leaflet)

opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")

```


```{r, include=FALSE, echo=FALSE}
academic15 <- read_excel("academic15.xlsx")
academic15 <- academic15 |> 
  select(c("Board Number", "Board Name", "School Number", "School Name", "Street", "Latitude", "Longitude", 
           "Percentage of Grade 6 Students Achieving the Provincial Standard in Reading", "School Level",
           "Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))
academic15 <- academic15 |> 
  mutate('Start Year' = 2015)


### 16
academic16 <- read_excel("academic16.xlsx")
academic16 <- academic16 |> 
  select(c("Board Number", "Board Name", "School Number", "School Name", "Street", "Latitude", "Longitude", 
           "Percentage of Grade 6 Students Achieving the Provincial Standard in Reading", "School Level",
           "Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))
academic16 <- academic16 |> 
  mutate('Start Year' = 2016)

### 17
academic17 <- read_excel("academic17.xlsx")
academic17 <- academic17 |> 
  select(c("Board Number", "Board Name", "School Number", "School Name", "Street", "Latitude", "Longitude", 
           "Percentage of Grade 6 Students Achieving the Provincial Standard in Reading", "School Level",
           "Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))
academic17 <- academic17 |> 
  mutate('Start Year' = 2017)

### 18
academic18 <- read_excel("academic18.xlsx")
academic18 <- academic18 |> 
  select(c("Board Number", "Board Name", "School Number", "School Name", "Street", "Latitude", "Longitude", 
           "Percentage of Grade 6 Students Achieving the Provincial Standard in Reading", "School Level",
           "Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))
academic18 <- academic18 |> 
  mutate('Start Year' = 2018)

### 19
academic19 <- read_excel("academic19.xlsx")
academic19 <- academic19 |> 
  select(c("Board Number", "Board Name", "School Number", "School Name", "Street", "Latitude", "Longitude", 
           "Percentage of Grade 6 Students Achieving the Provincial Standard in Reading", "School Level",
           "Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))
academic19 <- academic19 |> 
  mutate('Start Year' = 2019)

academic <- rbind(academic15, academic16, academic17, academic18, academic19)

elementary <- academic |> filter(`School Level` == "Elementary")
secondary <- academic |> filter(`School Level` == "Secondary")

# clean_academic <- na.omit(academic)

## elementary
clean_elementary <- elementary |> filter(`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading` != "NA")
clean_elementary <- na.omit(clean_elementary)

clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading` <- str_sub(clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading`, end = -2) 

clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading` <- as.numeric(clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading`)
clean_elementary <- na.omit(clean_elementary)
clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading` <- clean_elementary$`Percentage of Grade 6 Students Achieving the Provincial Standard in Reading` * 0.01 

clean_elementary <- clean_elementary |> select(-c("Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt"))


## secondary
clean_secondary <- secondary |> filter(`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt` != "NA")
clean_secondary <- na.omit(clean_secondary)

clean_secondary <- clean_secondary |> select(-c("Percentage of Grade 6 Students Achieving the Provincial Standard in Reading"))
clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt` <- str_sub(clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt`, end = -2)


clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt` <- as.numeric(clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt`)

clean_secondary <- na.omit(clean_secondary)
clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt` <- clean_secondary$`Percentage of Students That Passed the Grade 10 OSSLT on Their First Attempt` * 0.01 

```

```{r, include=FALSE, echo=FALSE}
school_info <- read_excel("school_info.xlsx")
school_info <- school_info |>
  select("Region",
         "Board Number",
         "Board Name",
         "Board Language",
         "School Number",
         "School Name",
         "School Level",
         "School Language",
         "School Type",
         "City")
```

```{r, include=FALSE, echo=FALSE}

data8 <- read_excel("gender_2015.xlsx")
data8 <- data8 %>%
  mutate('Start Year' = 2015)

data9 <- read_excel("gender_2016.xlsx")
data9 <- data9 %>%
  mutate('Start Year' = 2016)

data10 <- read_excel("enrolment_by_gender_1718_en.xlsx")
data10 <- data10 %>%
  mutate('Start Year' = 2017)

data11 <- read_excel("enrolment_by_gender_1819_en.xlsx")
data11 <- data11 %>%
  mutate('Start Year' = 2018)

data12 <- read_excel("gender_1920_en.xlsx")
data12 <- data12 %>%
  mutate('Start Year' = 2019)

data11 <- data11 |>
  rename(
    `Elementary Male Enrolment` = `Elementary Male`,
    `Elementary Female Enrolment` = `Elementary Female`,
    `Secondary Male Enrolment` = `Secondary Male`,
    `Secondary Female Enrolment` = `Secondary Female`,
    `Total Male Enrolment` = `Total Male`,
    `Total Female Enrolment` = `Total Female`
  )

data12 <- data12 |>
  rename(
    `Elementary Male Enrolment` = `Elementary Male`,
    `Elementary Female Enrolment` = `Elementary Female`,
    `Secondary Male Enrolment` = `Secondary Male`,
    `Secondary Female Enrolment` = `Secondary Female`,
    `Total Male Enrolment` = `Total Male`,
    `Total Female Enrolment` = `Total Female`
  )

student <- rbind(data8, data9, data10, data11, data12)
```

```{r, include=FALSE, echo=FALSE}
data15 <- read_excel("edu_gender_1516_en.xlsx")
data15 <- data15 %>%
  mutate('Start Year' = 2015)

data16 <- read_excel("edu_gender_1617_en_0.xlsx")
data16 <- data16 %>%
  mutate('Start Year' = 2016)

data17 <- read_excel("edu_gender_1718.xlsx")
data17 <- data17 %>%
  mutate('Start Year' = 2017)

data18 <- read_excel("edu_gender_1819.xlsx")
data18 <- data18 %>%
  mutate('Start Year' = 2018)

data19 <- read_excel("edu_gender_1920.xlsx")
data19 <- data19 %>%
  mutate('Start Year' = 2019)

educator <- rbind(data15, data16, data17, data18, data19) |>
  select(-c("Academic Year"))


```

```{r, include=FALSE, echo=FALSE}
expulsion <- read_excel("expulsions_0708_2021.xlsx")
expulsion <- expulsion |>
  select("Board Number",
         "Board Name",
         "2007-2008 Expulsion Rate",
         "2008-2009 Expulsion Rate",
         "2009-2010 Expulsion Rate",
         "2010-2011 Expulsion Rate",
         "2011-2012 Expulsion Rate",
         "2012-2013 Expulsion Rate",
         "2013-2014 Expulsion Rate",
         "2014-2015 Expulsion Rate",
         "2015-2016 Expulsion Rate",
         "2016-2017 Expulsion Rate",
         "2017-2018 Expulsion Rate",
         "2018-2019 Expulsion Rate",
         "2019-2020 Expulsion Rate",
         "2020-2021 Expulsion Rate")

expulsion <- expulsion |>
  mutate("2007" = 2007,
         "2008" = 2008,
         "2009" = 2009,
         "2010" = 2010,
         "2011" = 2011,
         "2012" = 2012,
         "2013" = 2013,
         "2014" = 2014,
         "2015" = 2015,
         "2016" = 2016,
         "2017" = 2017,
         "2018" = 2018,
         "2019" = 2019,
         "2020" = 2020)

new_expulsion <- data.frame("Expulsion Rate" = c(
         expulsion$"2007-2008 Expulsion Rate",
         expulsion$"2008-2009 Expulsion Rate",
         expulsion$"2009-2010 Expulsion Rate",
         expulsion$"2010-2011 Expulsion Rate",
         expulsion$"2011-2012 Expulsion Rate",
         expulsion$"2012-2013 Expulsion Rate",
         expulsion$"2013-2014 Expulsion Rate",
         expulsion$"2014-2015 Expulsion Rate",
         expulsion$"2015-2016 Expulsion Rate",
         expulsion$"2016-2017 Expulsion Rate",
         expulsion$"2017-2018 Expulsion Rate",
         expulsion$"2018-2019 Expulsion Rate",
         expulsion$"2019-2020 Expulsion Rate",
         expulsion$"2020-2021 Expulsion Rate"),
         "Start Year" = c(
           expulsion$"2007",
         expulsion$"2008",
         expulsion$"2009",
         expulsion$"2010",
         expulsion$"2011",
         expulsion$"2012",
         expulsion$"2013",
         expulsion$"2014",
         expulsion$"2015",
         expulsion$"2016",
         expulsion$"2017",
         expulsion$"2018",
         expulsion$"2019",
         expulsion$"2020"
         ),
         "Board Name" = c(
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name",
           expulsion$"Board Name"
         ),
         "Board Number" = c(
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number",
           expulsion$"Board Number"
         ))
```

```{r, include=FALSE, echo=FALSE}
suspension <- read_excel("suspensions_0708-2021.xlsx")

suspension <- suspension |>
  select("Board Number",
         "Board Name",
         "2007-2008 Suspension Rate",
         "2008-2009 Suspension Rate",
         "2009-2010 Suspension Rate",
         "2010-2011 Suspension Rate",
         "2011-2012 Suspension Rate",
         "2012-2013 Suspension Rate",
         "2013-2014 Suspension Rate",
         "2014-2015 Suspension Rate",
         "2015-2016 Suspension Rate",
         "2016-2017 Suspension Rate",
         "2017-2018 Suspension Rate",
         "2018-2019 Suspension Rate",
         "2019-2020 Suspension Rate",
         "2020-2021 Suspension Rate")

suspension <- suspension |>
  mutate("2007" = 2007,
         "2008" = 2008,
         "2009" = 2009,
         "2010" = 2010,
         "2011" = 2011,
         "2012" = 2012,
         "2013" = 2013,
         "2014" = 2014,
         "2015" = 2015,
         "2016" = 2016,
         "2017" = 2017,
         "2018" = 2018,
         "2019" = 2019,
         "2020" = 2020)

new_suspension <- data.frame("Suspension Rate" = c(
         suspension$"2007-2008 Suspension Rate",
         suspension$"2008-2009 Suspension Rate",
         suspension$"2009-2010 Suspension Rate",
         suspension$"2010-2011 Suspension Rate",
         suspension$"2011-2012 Suspension Rate",
         suspension$"2012-2013 Suspension Rate",
         suspension$"2013-2014 Suspension Rate",
         suspension$"2014-2015 Suspension Rate",
         suspension$"2015-2016 Suspension Rate",
         suspension$"2016-2017 Suspension Rate",
         suspension$"2017-2018 Suspension Rate",
         suspension$"2018-2019 Suspension Rate",
         suspension$"2019-2020 Suspension Rate",
         suspension$"2020-2021 Suspension Rate"),
         "Start Year" = c(
           suspension$"2007",
         suspension$"2008",
         suspension$"2009",
         suspension$"2010",
         suspension$"2011",
         suspension$"2012",
         suspension$"2013",
         suspension$"2014",
         suspension$"2015",
         suspension$"2016",
         suspension$"2017",
         suspension$"2018",
         suspension$"2019",
         suspension$"2020"
         ),
         "Board Name" = c(
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name",
           suspension$"Board Name"
         ),
         "Board Number" = c(
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number",
           suspension$"Board Number"
         ))

# educator 
# student 
# academic c
# school_info c
# new expulsion c
# new suspension c

new_expulsion <- new_expulsion |> filter(2015 <= Start.Year) |> filter(Start.Year <= 2019)

new_suspension <- new_suspension |> filter(2015 <= Start.Year) |> filter(Start.Year <= 2019)

```


```{r, include = FALSE, echo=FALSE}

# merge for elementary
dat1 <- merge(
  # Data
  x = clean_elementary,
  y = educator,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat1 <- dat1 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Start Year.x` == `Start Year.y`) |>
  select(-c(`Board Name.y`, `Start Year.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Start Year" = `Start Year.x`)

dat1 <- dat1 |> select(-c("Secondary Male Educators", "Secondary Female Educators", "Total Male Educators", "Total Female Educators", "Total Educators"))
```

```{r, include = FALSE, echo=FALSE}
dat1 <- merge(
  # Data
  x = dat1,
  y = student,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat1 <- dat1 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Start Year.x` == `Start Year.y`) |>
  select(-c(`Board Name.y`, `Start Year.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Start Year" = `Start Year.x`)

dat1 <- dat1 |> select(-c("Secondary Male Enrolment", "Secondary Female Enrolment", "Total Male Enrolment", "Total Female Enrolment"))
```

```{r, include = FALSE, echo=FALSE}
school_info <- school_info |> rename("Geo-Region" = "Region")

dat1 <- merge(
  # Data
  x = dat1,
  y = school_info,
  # List of variables to match
  by.x = "School Number",
  by.y = "School Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat1 <- dat1 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Board Number.x` == `Board Number.y`,
         `School Name.x` == `School Name.y`,
         `School Level.x` == `School Level.y`,
         `Board Language.x` == `Board Language.y`) |>
  select(-c(`Board Name.y`, `Board Language.y`, `School Level.y`, `School Name.y`, `Board Number.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Board Number" = `Board Number.x`,
         "School Name" = `School Name.x`,
         "School Level" = `School Level.x`,
         "Board Language" = `Board Language.x`
         )

```

```{r, include = FALSE, echo=FALSE}
dat1 <- merge(
  # Data
  x = dat1,
  y = new_expulsion,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board.Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat1 <- dat1 |>
  filter(`Start Year` == `Start.Year`) |>
  select(-c(`Start.Year`, `Board.Name`))

```

```{r, include = FALSE, echo=FALSE}

dat1 <- merge(
  # Data
  x = dat1,
  y = new_suspension,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board.Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat1 <- dat1 |>
  filter(`Start Year` == `Start.Year`) |>
  select(-c(`Start.Year`, `Board.Name`)) |>
  rename("Expulsion Rate" = `Expulsion.Rate`,
         "Suspension Rate" = `Suspension.Rate`)

```

```{r}
## merge for secondary
dat2 <- merge(
  # Data
  x = clean_secondary,
  y = educator,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat2 <- dat2 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Start Year.x` == `Start Year.y`) |>
  select(-c(`Board Name.y`, `Start Year.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Start Year" = `Start Year.x`)

dat2 <- dat2 |> select(-c("Elementary Male Educators", "Elementary Female Educators", "Total Male Educators", "Total Female Educators", "Total Educators"))


dat2 <- merge(
  # Data
  x = dat2,
  y = student,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat2 <- dat2 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Start Year.x` == `Start Year.y`) |>
  select(-c(`Board Name.y`, `Start Year.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Start Year" = `Start Year.x`)

dat2 <- dat2 |> select(-c("Elementary Male Enrolment", "Elementary Female Enrolment", "Total Male Enrolment", "Total Female Enrolment"))


dat2 <- merge(
  # Data
  x = dat2,
  y = school_info,
  # List of variables to match
  by.x = "School Number",
  by.y = "School Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat2 <- dat2 |>
  filter(`Board Name.x` == `Board Name.y`,
         `Board Number.x` == `Board Number.y`,
         `School Name.x` == `School Name.y`,
         `School Level.x` == `School Level.y`,
         `Board Language.x` == `Board Language.y`) |>
  select(-c(`Board Name.y`, `Board Language.y`, `School Level.y`, `School Name.y`, `Board Number.y`)) |>
  rename("Board Name" = `Board Name.x`,
         "Board Number" = `Board Number.x`,
         "School Name" = `School Name.x`,
         "School Level" = `School Level.x`,
         "Board Language" = `Board Language.x`
         )

dat2 <- merge(
  # Data
  x = dat2,
  y = new_expulsion,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board.Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat2 <- dat2 |>
  filter(`Start Year` == `Start.Year`) |>
  select(-c(`Start.Year`, `Board.Name`))

dat2 <- merge(
  # Data
  x = dat2,
  y = new_suspension,
  # List of variables to match
  by.x = "Board Number",
  by.y = "Board.Number",
  # Which obs to keep?
  all.x = FALSE,
  all.y = FALSE
)

dat2 <- dat2 |>
  filter(`Start Year` == `Start.Year`) |>
  select(-c(`Start.Year`, `Board.Name`)) |>
  rename("Expulsion Rate" = `Expulsion.Rate`,
         "Suspension Rate" = `Suspension.Rate`)
```


```{r}
dat1$`Elementary Male Educators` <- as.numeric(dat1$`Elementary Male Educators`)
dat1$`Elementary Male Enrolment` <- as.numeric(dat1$`Elementary Male Enrolment`)
dat1$`Elementary Female Enrolment` <- as.numeric(dat1$`Elementary Female Enrolment`)
dat1 <- na.omit(dat1)

dat2$`Secondary Male Educators` <- as.numeric(dat2$`Secondary Male Educators`)
dat2$`Secondary Female Educators` <- as.numeric(dat2$`Secondary Female Educators`)
dat2$`Secondary Male Enrolment` <- as.numeric(dat2$`Secondary Male Enrolment`)
dat2$`Secondary Female Enrolment` <- as.numeric(dat2$`Secondary Female Enrolment`)
dat2 <- na.omit(dat2)
dat2 <- dat2 |> filter(Latitude >= -76)

```



```{r, include=FALSE, echo=FALSE}
# create new variable student teacher ratio


dat1 <- dat1 |>
  mutate("Elementary Student Teacher Ratio" = (`Elementary Male Enrolment` + `Elementary Female Enrolment`) /
           (`Elementary Male Educators` + `Elementary Female Educators`))

dat2 <- dat2 |> 
  mutate("Secondary Student Teacher Ratio" = (`Secondary Male Enrolment` + `Secondary Female Enrolment`) / 
           (`Secondary Male Educators` + `Secondary Female Educators`))

```


# Introduction

Education is one of the fundamental building blocks of any society and plays a vital role in shaping the future of individuals, communities, and nations. Among all years in school, k-12 education has the most significant influence on students because their personality and vision of the world are primarily shaped in elementary and secondary schools. Factors such as the school's suspension rate, drop rate, number of students, full-time educators, and geographic locations can partially reflect a school's academic performance. For example, suspension rates in primary and secondary schools are of concern as they are associated with adverse outcomes such as reduced academic performance, increased dropout rates, and a higher likelihood of involvement in delinquent behavior. On the other hand, other factors like the teacher-student ratio can suggest how many school resources and attention teachers each student can get. Thus a smaller ratio might imply a higher academic performance of a school board. Hence, from these factors, we can analyze their impacts on a school's academic performance, and understanding the relationship between them is crucial for policymakers, educators, and stakeholders in designing effective strategies to improve educational outcomes.

With this in mind, I want to investigate the following question: What factors of a school will influence their academic performance and their relationships? To be more specific, what are the relationships between each school board's academic performance and their suspension rate, drop rate, number of students, number of full-time educators, and geographic locations?

The data I obtained to answer the questions is from 
[Ontario public data website](https://www.ontario.ca/page/government-ontario) and 
[Open Canada](https://www.canada.ca/en.html) searching with keyword _school board_. The datasets contain information about all public schools, including elementary and secondary schools, in Ontario. All datasets I can find are from 2015 to 2019, so I will also limit my scope to this range. 

## Research Question
What factors of a school will influence their academic performance and their relationships? To be more specific, what are the relationships between each school board's academic performance and their suspension rate, drop rate, number of students, number of full-time educators, and geographic locations?


# Methods
## Data Collection

We obtain all our datasets from [Ontario public data website](https://www.ontario.ca/page/government-ontario) and [Open Canada](https://www.canada.ca/en.html) by searching for the keyword `school board`. To be more specific, we download datasets for information on [each school board's academic standing](https://open.canada.ca/data/en/dataset/d85f68c5-fcb0-4b4d-aec5-3047db47dcd5), [each school board's full-time educators](https://data.ontario.ca/dataset/number-of-full-time-equivalent-educators-by-gender-and-school-board-in-ontario), [each school board's students](https://data.ontario.ca/en/dataset/school-enrolment-by-gender), [each school board's schools](https://data.ontario.ca/en/dataset/ontario-public-school-contact-information), [each school board's expulsion rate](https://data.ontario.ca/en/dataset/expulsion-rates-by-school-board), and [each school board's suspension rate](https://data.ontario.ca/en/dataset/suspension-rates-by-school-board).

We download all the datasets mentioned above and load than using `read_excel`. Since datasets for academic performance, educators, and students have their data stored in a separate file for different years, we simply use `rbind` to combine them into a single table. In addition, because the datasets of expulsion and suspension rates have a format that each roll is a school board and each column is the suspension rate for a specific year, we created two new datasets to make them into the correctly formatted datasets.

## Data Wrangling
Since the original datasets collected are too messy to summarize and display, I will perform data cleaning and wrangling first and then show the cleaned form of the dataset.

### Data Cleaning
We drop all the irrelevant variables that do not provide insight into the question of interest, such as variable `phone number` in the dataset of schools information, variable `Four Year Graduation Rate` in the dataset for academic performance, and some categorical variables in datasets for expulsion rate and suspension rate. We also notice that for enrollment status datasets, despite all containing the same information, the column names are different across datasets, and thus we rename them to the same name for future simplicity. Next, since the academic performance datasets contain information for both elementary and secondary schools identified by the variable `School Level`, I separate the single dataset into two dataframes, one for elementary and another for secondary. 

In addition, the passing rates for EQAO and OSSLT are documented as characters with the format `x%`, and the elementary male and female enrollment numbers and secondary male and female enrollment numbers are as characters even though they are numbers. Hence, I convert them to numerical variables, which are more intuitive.

### Data Merging
For elementary and secondary datasets each, we merge them with other five tables to get two final datasets that will be used in the analysis. During the merging, we ensure common variables `Board Number`, `Start Year`, `School Name` are matched for each row every time we merge two datasets. After taking a closer look at the two merged datasets, we detected that elementary contains 5 missing values and secondary contains 8 missing values, so we dropped all rows with missing values. 


### Feature Engineering
We want to create two variables `Elementary Student-Teacher Ratio` and `Secondary Student-Teacher Ratio`, which is the number of students in the school board divided by the number of teachers in the row. We create this using function `mutate` and the two new variables are calculated as previously mentioned. Now, we obtain the final datasets that we can work with. The elementary school information dataset has 15,003 rows and 24 columns, and the secondary school information dataset has 3414 rows and 24 variables.




## Data Summary

Table below is a description of the crucial variables that we used in this report. We only include part of the variables since others such as `Latitude` and `Longitude` are just for visualization not analysis.

```{r, echo=FALSE}
data <- read.csv("document.csv")

data <- data |>
  select(-("ID")) |>
  rename("Variable Name" = Variable.Name,
         "Variable Description" = Occupational.title)

knitr::kable(data)
```

## Data Visualization

```{r}
plt1 <- dat1 |>
  filter(`Start Year` == 2018) |>
  ggplot() +
  geom_bar(aes(x = `Geo-Region`, fill = `Geo-Region`)) +
  labs(title = "Figure 1 - A barplot of the number of schools in each region",
       y = "Count",
       x = "School Region") +
  coord_flip() +
  theme_bw()

plt2 <- dat2 |>
  filter(`Start Year` == 2018) |>
  ggplot() +
  geom_bar(aes(x = `Geo-Region`, fill = `Geo-Region`)) +
  labs(title = "Figure 1 - A barplot of the number of schools in each region",
       y = "Count",
       x = "School Region") +
  coord_flip() +
  theme_bw()

grid.arrange(plt1, plt2, ncol=1)
```


## Tools Used

The tools used in data wrangling are from packages `tidyverse` and `dplyr`. All tables are created with `kable`. Figures and plots are made using `ggplot2`, interactive visuals use package `plotly`, and maps are created with leaflet. 

# Results
## Summary Visuals
## Modelling
## Classification

# Conclusion